---
title: "X. Intro to Data Wrangling using R"
subtitle: "NACC Workshop"
author: "Mansour Abdoli, PhD"
date: "February 11, 2025"
format: revealjs
editor: source
---

## Statistical Process

::: incremental
::: columns
::: {.column style="font-size: 85%"}
-   Question
    -   Parameter(s) of interest <br> ($\pi$, $\mu$, $\rho$)
    -   Response/Explanatory variable(s)
:::

::: {.column style="font-size: 85%"}
-   Data
    - Sampling or Experiment Design
    - Collecting Data
    - **Data Wrangling**
:::
:::
:::

::: incremental
::: columns
::: {.column style="font-size: 85%"}
-   Descriptive Analysis
    -   **Visualizing**
    -   Summarizing
:::

::: {.column style="font-size: 85%"}
-   Inferential Analysis
    -   **Modeling**
    -   Generalizing and Predicting
:::
:::
:::

## Inference on Single Variable: Variation

- Single Variable (Response)
  - Sample Statistic vs Expected Value.
  - Inference on the Expected Value.

```{r}
set.seed(123)
x <- rnorm(100)
y <- rnorm(100, sd = 2)
z=3*x+rnorm(100,sd=1)

par(mar = c(1,1,1,1), oma = c(1,0,0,0), mgp=c(1,.5,0), mfrow=c(1, 2))
hist(x, axes=FALSE, ann=F, xlim = c(-5,5))
hist(y, axes=FALSE, ann=F, xlim = c(-5,5))
mtext(text = 'Which one gives a better estimate?',
      side = 1, line = 0, adj = 0.5, outer = T)
```
## Inference on Single Variable: Variation

- Single Variable (Response)
  - Sample Statistic vs Expected Value.
  - Inference on the Expected Value.

```{r}
set.seed(123)
k <- sample.int(2, 100, replace = TRUE, prob = c(0.5, .5))
j <- sample.int(2, 100, replace = TRUE, prob = c(0.1, .9))

par(mar = c(1,1,1,1), oma = c(1,0,0,0), mgp=c(1,.5,0), mfrow=c(1, 2))
barplot(c(sum(k==1), sum(k==2)), axes=FALSE, ann=F, xlim = c(-1,3))
barplot(c(sum(j==1), sum(j==2)), axes=FALSE, ann=F, xlim = c(-1,3))
mtext(text = 'Which generates less variable samples?',
      side = 1, line = 0, adj = 0.5, outer = T)
```

## Inference on Association
- Expected Value of Response vs Explanatory Values.
- Inference on the Expected Value
- Inference on the Association Parameter(s)

```{r}
par(mar = c(1,1,1,1), oma = c(1,0,0,0), mgp=c(1,.5,0), mfrow=c(1, 2))
plot(x, y, axes=FALSE, ann=F, xlim = c(-3,3), ylim = c(-8,8))
plot(x, z, axes=FALSE, ann=F, xlim = c(-3,3), ylim = c(-8,8))
mtext(text = 'Which one shows an association?',
      side = 1, line = 0, adj = 0.5, outer = T)

```

  
## Reading Local & Remote Files:

```{r, eval=FALSE}
alz <- read.csv('https://drive.google.com/uc?export=download&id=1e_mA_i9-7dyGAb-SpeIS1OvimMi9HKU1')
uds <- read.csv('https://drive.google.com/uc?export=download&id=1YQxENus9dc2k42-NPbngvH6qogzEyDqj')
mri <- read.csv('https://drive.google.com/uc?export=download&id=10dJUNf1esACoP22BxMI9XI3r34ZDuu09')
```


